\bf A Comparision of Bootstrap Confidence Interval Methods

### 2022-12-09 project/PaperTemplate/paper.tex
Finished paper

  \author{
        Justin Papagelis \thanks{The authors gratefully acknowledge '\,'
Profossor Wagaman} \\
    Department of Mathematics and Statistics, Amherst College\\
      }
  \maketitle
} \fi

@@

Bootstrapping is a statistical method of resampling that allows the
estimation of a test statistic from an unknown distribution. In
particular, bootstrapping is a computational heavy method which is
useful for many different situations.

First, we introduce the non-parametric bootstrap. Suppose we have a
random sample \(X = (x_1,x_2,\dots,x_n)\) from our unknown distribution,
\(F\) and a statistic of interest, \(\hat{\theta} = \hat{\theta}(X)\)
\citep{EfronCasi}. Ideally, the desired test statistic could be found by
repeatedly sampling new reproductions of \(X\) from \(F\). However \(F\)
is unknown, so this is not possible. The non-parametric bootstrap
creates an estimate \(\hat{F}\) from \(F\) using our sample, \(X\),
without making any parametric assumptions about \(F\) (such as its
distribution type). Therefore, the bootstrap sample could be represented
as \(X^* = (x^*_1, x^*_2, \dots, x^*_n)\) where each \(x^*_i\) is
sampled randomly with equal probability and with replacement from our
original sample: \(\{x_1,x_2,\dots,x_n\}\). From this bootstrap sample,
a bootstrap replication of the test statistic can be computed using
\(\hat{\theta}^* = \hat{\theta}(X^*)\). A large number, \(B\), of
bootstrap samples are drawn independently and the corresponding
bootstrap replication of the test statistic is calculated.
\[\hat{\theta}^{*b} = \hat{\theta}(X^{*b}) \text{ for } b = 1,2, \dots, B.\]
The bootstrap estimate of the test statistic is the empirical value of
the test statistic from all of the \(\hat{\theta}^{*b}\) replications.
As \(B\) increases, \(\hat{F}\) approaches \(F\) which means that the
test statistic of interest approaches its true value as well.

@@

We demonstrate performing a non-parametric bootstrap below using
\texttt{SnowGR} which gives the official snowfall dataset by month for
Grand Rapids, MI, starting in 1893. We will be using the \texttt{Dec}
variable which is the number of inches of snow that fell in December of
each year. In particular, we are interested in the sample mean and later
finding confidence intervals for population mean of \texttt{Dec}. The
distribution is below (Figure 1).

@@

Below, in Figure 2, is the histogram of bootstrapped values. Since we
performed a large number of bootstrap replications, the bootstrapped
distribution appears to be approximately Normal.

@@

Confidence intervals are tools that are used to estimate a parameter.
Specifically, a confidence interval gives a range of values in which the
true value of the parameter may lie. An \(\alpha\)-level standard
confidence interval is given by
\[\hat{\theta}_S[\alpha] = \hat{\theta} \pm z_{\alpha}\hat{\sigma},\]
where \(\hat{\theta}\) is a point estimate of the parameter of interest
\(\theta\), \(\hat{\sigma}\) is the estimate of the standard deviation
of \(\hat{\theta}\) and \(z_{\alpha}\) is the \((100 *\alpha)\)th
percentile of the normal deviation \citep{Efron86}. We say that the
confidence interval constructed in this manner has a chance of capturing
the true parameter with a probability of \(\alpha\).

The standard confidence interval is built based on the assumption that
the distribution from which we are sampling is Normal. As such, the
standard confidence interval is sometimes called the normal confidence
interval for a bootstrapped sample. This means that for an unknown
skewed distribution, the standard confidence interval could present an
incorrect range. However, the same process can be used with bootstrap
sampling to form the bootstrap percentile method. This means that an
approximate bootstrap confidence interval will be created in the same
automatic way that the standard confidence interval was created. For
bootstrapped confidence intervals, the number of bootstrap replications
\(B\) must be large (around 2,000) due to the nature of confidence
intervals requiring greater accuracy \citep{Efron86}.

@@

This means that we can say we are 95\% confident that the true mean
number of inches of snow that fall in December is betwwen 13.86 inches
and 17.66 inches. Our original histogram with the confidence interval is
below (Figure 3).

@@

The percentile method interval is defined as the interval between the
\(100 * \alpha\) and the \(100(1 - \alpha)\) percentiles of the
bootstrap distribution of \(\hat{\theta}\). That is, the
\((1 - 2\alpha)\) coverage interval can be defined as
\([\hat{\theta}^*_\alpha,\hat{\theta}^*_{1-\alpha}]\)
\citep[\citet{EfronCasi}]{Efron86}. To go further, we can define
\(\hat{G}(t)\) as the bootstrap cdf, or the proportion of bootstrap
samples less than \(t\):
\[\hat{G}(t) = \frac{\#\{\hat{\theta}^{*b} \leq t\}}{B}.\] Thus the
\(\alpha\)th percentile point of the distribution is given by
\[\hat{\theta}_p[\alpha] = \hat{\theta^*_\alpha} = \hat{G}^{-1}(\alpha).\]
It follows that the percentile interval can be represented as
\[\left [ \hat{G}^{-1}(\alpha),\hat{G}^{-1}(1-\alpha) \right ].\] In the
case that the bootstrap distribution of
\(\hat{\theta}^* \sim N(\hat{\theta}, \hat{\sigma}^2)\), the
corresponding percentile interval would be equivalent to the standard
interval. However, this is not usually the case. When the bootstrap
distribution is non-normal, we can suppose that there exists, for all
\(\theta\), \[\hat{\phi} \sim N(\phi, \tau^2),\] for some monotone
transformation \(\hat{\phi} = g(\hat{\theta}), \phi = g(\theta)\), and
\(\tau\) is a constant. In other words, this transformation perfectly
normalizes the distribution of \(\hat{\theta}\). This transformation
invariant can be applied to the bootstrap replications such that
\[\hat{\phi}^{*b} = g\left( \hat{\theta}^{*b}\right ) \text{ for } b = 1,2,\dots, B.\]
The corresponding percentiles of the distribution transform similarly,
\(\hat{\phi}^*_\alpha = g \left ( \hat{\theta}^*_\alpha \right )\). Or
we can say that the \((1 - 2\alpha)\) percentile interval is
\(\hat{\phi} \pm \tau z_\alpha\) which can also be represented as
\([\hat{\phi}^*_\alpha,\hat{\phi}^*_{1-\alpha}]\). This means that the
interval on the \(\theta\) scale can be defined as
\[\hat{\theta}^*_\alpha = g^{-1}(\hat{\phi} \pm \tau z_\alpha).\] This
also can be represented as an interval,
\[\left [ g^{-1}(\hat{\phi} \pm \tau z_{1-\alpha}), g^{-1}(\hat{\phi} \pm \tau z_\alpha) \right ].\]
Therefore, the percentile method produces a correct interval for
\(\phi\) and due to the transformation invariance, also produces a
correct percentile interval for \(\theta\). This method assumes the
existence of some monotone normalizing mapping
\(\hat{\phi} = g(\hat{\theta}), \phi = g(\theta)\) and relies on that to
create a correct interval. Since the process is automatic, we do not
need to know the transformation itself, only that it exists. However, in
some cases, no monotone normalizing mapping will exist \citep{Efron86}.

@@

Therefore, we can say that we are 95\% confident that the true
population mean of inches of snow that fall in December is between 13.91
inches and 17.72 inches. Our original histogram with the percentile
confidence interval is shown below (Figure 4).

@@

The next method we will be looking at is the bias-corrected percentile
method (BC method) which is an improvement upon the previous percentile
method because we now take into account the possibility of bias. It can
be shown that \(\hat{\theta}\) is biased upwards relative to \(\theta\)
which means that the confidence intervals should be adjusted downwards
\citep[\citet{EfronCasi}]{Efron86}. From our simulated bootstrap
replications
\(\hat{\theta}^{*1}, \hat{\theta}^{*2}, \dots ,\hat{\theta}^{*B},\)
define \[p_0 = \frac{\#\{\hat{\theta}^{*b} \leq \theta\}}{B},\] and
define the bias-correction value \[z_0 = \Phi^{-1}(p_0),\] where
\(\Phi^{-1}\) is the inverse function of the standard normal cdf. Thus,
we define a transformation
\(\hat{\phi} = g(\hat{\theta)}, \phi = g(\theta)\) such that for any
\(\theta\), \[\hat{\phi} \sim N(\phi - z_0\tau, \tau^2),\] with \(z_0\)
and \(\tau\) constants. This means that we can say the bias-corrected
method has an \(\alpha\)-level endpoint which can be represented as
\[\hat{\theta}_{BC}[\alpha] = \hat{G}^{-1} \left [ \Phi \left ( 2z_0 + z_\alpha\right ) \right ].\]
If \(\hat{G} = 0.50\), then half of the bootstrap distribution is less
than \(\hat{\theta}\) and our bias-correction value \(z_0 = 0\). In this
case, the confidence interval produced by BC would be the same interval
that was produced by the percentile method.

@@

Using this method to create a 95\% confidence interval has a couple more
steps because we need to find the bias-correction value. First we
calculate the sample mean:

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{alphalow }\OtherTok{\textless{}{-}} \FloatTok{0.025}\NormalTok{; alphahigh }\OtherTok{\textless{}{-}} \FloatTok{0.975}

\NormalTok{newlow }\OtherTok{\textless{}{-}} \FunctionTok{pnorm}\NormalTok{(}\DecValTok{2}\SpecialCharTok{*}\NormalTok{z0 }\SpecialCharTok{+} \FunctionTok{qnorm}\NormalTok{(alphalow)); }\CommentTok{\#newlow}
\NormalTok{newhigh }\OtherTok{\textless{}{-}} \FunctionTok{pnorm}\NormalTok{(}\DecValTok{2}\SpecialCharTok{*}\NormalTok{z0 }\SpecialCharTok{+} \FunctionTok{qnorm}\NormalTok{(alphahigh)); }\CommentTok{\#newhigh}

\FunctionTok{qdata}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{ mean, }\FunctionTok{c}\NormalTok{(newlow, newhigh), }\AttributeTok{data =}\NormalTok{ dec\_means)}
\end{Highlighting}
\end{Shaded}

@@

Therefore, we are 95\% confident that the true mean inches of snow that
fall in December in Grand Rapids, Michigan is between 14.01 inches and
17.82 inches. Our histogram (Figure 5) is shown below.

@@

A further modification upon the BC interval is the bias-corrected and
accelerated method (BCa). For this method, we do not assume the the
standard error, \(\tau\), is constant as we did in the BC interval
\citep[\citet{EfronCasi}]{Efron86}. Rather, we assume the existence of a
monotone transformation
\(\hat{\phi} = g(\hat{\theta)}, \phi = g(\theta)\) such that for any
\(\theta\),
\[\hat{\phi} \sim N(\phi - z_0\tau_\phi, \tau_\phi^2) \text{ where } \tau_\phi = 1+ a\phi.\]
The \(a\) is known as the acceleration and is a constant that describes
how the standard deviation of \(\hat{\phi}\) varies with \(\phi\). In
other words, \(a\) is proportional to the skewness of the bootstrap
distribution. For example, \(a=z_0\) for one-parameter exponential
families, however, there are many different algorithms to compute and
estimate \(a\) \citep{Flowers18}. Now, our \(\alpha\)-level endpoint for
a BCa confidence interval is
\[\hat{\theta}_{BCa}[\alpha] = \hat{G}^{-1} \left [ \Phi \left ( z_0 + \frac{z_0+z_\alpha}{1-a(z_0+z_a)} \right ) \right ].\]
We can observe that if \(a = 0\), then
\(\hat{\theta}_{BCa}[\alpha] = \hat{\theta}_{BC}[\alpha].\) When
calculating a BCa confidence interval, the acceleration value, \(a\), is
not a function of the bootstrap distribution and must be calculated
separately. However the process is algorithmic and can be calculated
without too much work. As we saw, each of the three previous methods
(percentile, BC, and BCa) all build upon each other and have less
restrictive assumptions, but this means that computation increases as we
loosen assumptions.

@@

In practice, we can use the jackknife procedure to estimate the
acceleration value for this method. The jackknife procedure is a
different resampling technique that is often used to estimate the bias
and variance of a large population. Thus, we will run the jackknife
procedure:

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{jackmean }\OtherTok{\textless{}{-}} \FunctionTok{mean}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{ jackknife\_results}\SpecialCharTok{$}\NormalTok{jack.values); jackmean}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{estimated\_a }\OtherTok{\textless{}{-}}\NormalTok{ (}\DecValTok{1}\SpecialCharTok{/}\DecValTok{6}\NormalTok{)}\SpecialCharTok{*}\FunctionTok{sum}\NormalTok{((jackknife\_results}\SpecialCharTok{$}\NormalTok{jack.values }\SpecialCharTok{{-}}\NormalTok{ jackmean)}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{)}\SpecialCharTok{/}
\NormalTok{  (}\FunctionTok{sum}\NormalTok{((jackknife\_results}\SpecialCharTok{$}\NormalTok{jack.values }\SpecialCharTok{{-}}\NormalTok{ jackmean)}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{))}\SpecialCharTok{\^{}}\NormalTok{(}\FloatTok{1.5}\NormalTok{); estimated\_a}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{a\_newlow }\OtherTok{\textless{}{-}} \FunctionTok{pnorm}\NormalTok{(z0 }\SpecialCharTok{+}\NormalTok{ (z0 }\SpecialCharTok{+} \FunctionTok{qnorm}\NormalTok{(alphalow))}\SpecialCharTok{/}
\NormalTok{                    (}\DecValTok{1}\SpecialCharTok{{-}}\NormalTok{estimated\_a}\SpecialCharTok{*}\NormalTok{(z0 }\SpecialCharTok{+} \FunctionTok{qnorm}\NormalTok{(alphalow)))); }
\NormalTok{a\_newhigh }\OtherTok{\textless{}{-}} \FunctionTok{pnorm}\NormalTok{(z0 }\SpecialCharTok{+}\NormalTok{ (z0 }\SpecialCharTok{+} \FunctionTok{qnorm}\NormalTok{(alphahigh))}\SpecialCharTok{/}
\NormalTok{                     (}\DecValTok{1}\SpecialCharTok{{-}}\NormalTok{estimated\_a}\SpecialCharTok{*}\NormalTok{(z0 }\SpecialCharTok{+} \FunctionTok{qnorm}\NormalTok{(alphahigh)))); }
\FunctionTok{qdata}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{ mean, }\FunctionTok{c}\NormalTok{(a\_newlow, a\_newhigh), }\AttributeTok{data =}\NormalTok{ dec\_means)}
\end{Highlighting}
\end{Shaded}

@@

The final method we explain is the bootstrap-t interval (or the
studentized method). Recall from earlier, our sample, \(X\) from which
we can calculate an estimate (\(\hat{\theta}(X)\)) of the parameter of
interest \(\theta\) \citep[\citet{Puth15}]{Efron86}. We can also
estimate \(\hat{\sigma}(X)\) for the standard error of \(\theta\). We
can use these parameters to find the Student's \(t\)-statistic defined
as: \[T = \frac{\hat{\theta} - \theta}{\hat{\sigma}}.\] As such, the
\(\alpha\)th percentile point of a confidence interval of \(\theta\)
would be \(\hat{\theta} - \hat{\sigma}T_{(\alpha)}\) where
\(T_{(\alpha)}\) represents the \(\alpha\)th percentile of the
\(t\)-distribution, \(T\). Unfortunately, the percentiles of this
\(t\)-distribution are unknown in most cases, but we can use
bootstrapping to estimate these percentiles. To do this, we perform a
large number, \(B\), of bootstrap samples, from which we can find the
bootstrap replications of the parameter of interest,
\(\hat{\theta^*} = \hat{\theta}(X^*)\), and the standard error,
\(\hat{\sigma^*} = \hat{\sigma}(X^*)\). From these we can calculate a
t-statistic for each bootstrap sample:
\[T^* = \frac{\hat{\theta}^* - \hat{\theta}}{\hat{\sigma}^*}.\] Using a
large number of these bootstrap samples, we can estimate the percentiles
of the \(t\)-distribution such that:
\[\hat{T}_{(\alpha)} = B*\alpha\text{th ordered value of all the bootstrap relications of }T^*.\]
This means that for \(B = 2,000\) and \(\alpha = 0.90\), then
\(\hat{T}_{(\alpha)}\) is the 1,800th ordered point of all of the
bootstrap replications of \(T^*\). It follows that the \(\alpha\)th
studentized confidence interval endpoint can be given with
\[\hat{\theta}_T[\alpha] = \hat{\theta} - \hat{\sigma}T_{(\alpha)},\]
where we can estimate the standard error using
\[\hat{\sigma} = \frac{1-\hat{\theta}^2}{\sqrt{n}}.\] One of the main
factors why the studentized method is so popular is because we assume
that our bootstrapped statistic is pivotal which means that the
confidence interval does not depend on any other parameters. Instead, we
can calculate the appropriate confidence interval for the parameter of
interest specifically from the bootstrapped statistics
\citep[\citet{Puth15}]{Efron86}.

@@

So we are 95\% confident that the true mean number of inches of snow
that fall in December in Grand Rapids, Michigan is between 14.03 inches
and 18.02 inches. This would give us the following confidence interval
(Figure 7)

@@

\includegraphics{paper_files/figure-latex/output the graphs/tables-1.pdf}

>>>>

  \author{
        Justin Papagelis \thanks{The authors gratefully acknowledge
Profossor Wagaman} \\
    Department of Mathematics and Statistics, Amherst College\\
      }
  \maketitle
} \fi

@@

Bootstrapping is a statistical method of resampling that allows the
estimation of a test statistic from an unknown distribution. In
particular, bootstrapping is a heavy computational method useful for
many situations.

First, we introduce the non-parametric bootstrap. Suppose we have a
random sample \(X = (x_1,x_2,\dots,x_n)\) from our unknown distribution,
\(F\) and a statistic of interest, \(\hat{\theta} = \hat{\theta}(X)\)
\citep{EfronCasi}. Ideally, the desired test statistic could be found by
repeatedly sampling new reproductions of \(X\) from \(F\). However,
\(F\) is unknown, so this is not possible. The non-parametric bootstrap
creates an estimate \(\hat{F}\) from \(F\) using our sample, \(X\),
without making any parametric assumptions about \(F\) (such as its
distribution type). Therefore, the bootstrap sample could be represented
as \(X^* = (x^*_1, x^*_2, \dots, x^*_n)\) where each \(x^*_i\) is
sampled randomly with equal probability and with replacement from our
original sample: \(\{x_1,x_2,\dots,x_n\}\). From this bootstrap sample,
a bootstrap replication of the test statistic can be computed using
\(\hat{\theta}^* = \hat{\theta}(X^*)\). A large number, \(B\), of
bootstrap samples are drawn independently, and the corresponding
bootstrap replication of the test statistic is calculated.
\[\hat{\theta}^{*b} = \hat{\theta}(X^{*b}) \text{ for } b = 1,2, \dots, B.\]
The test statistic's bootstrap estimate is the test statistic's
empirical value from all of the \(\hat{\theta}^{*b}\) replications. As
\(B\) increases, \(\hat{F}\) approaches \(F\), which means that the test
statistic of interest approaches its true value as well.

@@

We demonstrate performing a non-parametric bootstrap below using
\texttt{SnowGR}, which gives the official snowfall dataset by the month
for Grand Rapids, MI, starting in 1893. We will be using the
\texttt{Dec} variable, which is the number of inches of snow that fell
in December of each year. In particular, we are interested in the sample
mean and later finding confidence intervals for the population mean of
\texttt{Dec}. The distribution is shown (Figure 1).

@@

Figure 2 shows the histogram of bootstrapped values. Since we performed
a large number of bootstrap replications, the bootstrapped distribution
appears to be approximately Normal.

@@

Confidence intervals are tools that are used to estimate a parameter.
Specifically, a confidence interval gives a range of values in which the
true value of the parameter may lie. An \(\alpha\)-level standard
confidence interval is given by
\[\hat{\theta}_S[\alpha] = \hat{\theta} \pm z_{\alpha}\hat{\sigma},\]
where \(\hat{\theta}\) is a point estimate of the parameter of interest
\(\theta\), \(\hat{\sigma}\) is the estimate of the standard deviation
of \(\hat{\theta}\) and \(z_{\alpha}\) is the \((100 *\alpha)\)th
percentile of the normal deviation \citep{Efron86}. We say that the
confidence interval constructed in this manner can capture the true
parameter with a probability of \(\alpha\).

The standard confidence interval is built based on the assumption that
the distribution we are sampling is Normal. The standard confidence
interval is sometimes called the normal confidence interval for a
bootstrapped sample. This means that the standard confidence interval
could present an incorrect range for an unknown skewed distribution.
However, the same process can be used with bootstrap sampling to form
the bootstrap percentile method. This way, an approximate bootstrap
confidence interval will be created in the same automatic way that the
standard confidence interval was created. For bootstrapped confidence
intervals, the number of bootstrap replications \(B\) must be large
(around 2,000) due to the nature of confidence intervals requiring
greater accuracy \citep{Efron86}.

@@

Therefore, we are 95\% confident that the true mean number of inches of
snow that fall in December is between 13.86 inches and 17.66 inches. Our
original histogram with the confidence interval is shown (Figure 3).

@@

The percentile method interval is defined as the interval between the
\(100 * \alpha\) and the \(100(1 - \alpha)\) percentiles of the
bootstrap distribution of \(\hat{\theta}\). That is, the
\((1 - 2\alpha)\) coverage interval can be defined as
\([\hat{\theta}^*_\alpha,\hat{\theta}^*_{1-\alpha}]\)
\citep[\citet{EfronCasi}]{Efron86}. To go further, we can define
\(\hat{G}(t)\) as the bootstrap cdf, or the proportion of bootstrap
samples less than \(t\):
\[\hat{G}(t) = \frac{\#\{\hat{\theta}^{*b} \leq t\}}{B}.\] Thus the
\(\alpha\)th percentile point of the distribution is given by
\[\hat{\theta}_p[\alpha] = \hat{\theta^*_\alpha} = \hat{G}^{-1}(\alpha).\]
It follows that the percentile interval can be represented as
\[\left [ \hat{G}^{-1}(\alpha),\hat{G}^{-1}(1-\alpha) \right ].\] In the
case that the bootstrap distribution of
\(\hat{\theta}^* \sim N(\hat{\theta}, \hat{\sigma}^2)\), the
corresponding percentile interval would be equivalent to the standard
interval. However, this is only sometimes the case. When the bootstrap
distribution is non-normal, we can suppose that there exists, for all
\(\theta\), \[\hat{\phi} \sim N(\phi, \tau^2),\] for some monotone
transformation \(\hat{\phi} = g(\hat{\theta}), \phi = g(\theta)\), and
\(\tau\) is a constant. In other words, this transformation perfectly
normalizes the distribution of \(\hat{\theta}\). This transformation
invariant can be applied to the bootstrap replications such that
\[\hat{\phi}^{*b} = g\left( \hat{\theta}^{*b}\right ) \text{ for } b = 1,2,\dots, B.\]
The corresponding percentiles of the distribution transform similarly,
\(\hat{\phi}^*_\alpha = g \left ( \hat{\theta}^*_\alpha \right )\). Or
we can say that the \((1 - 2\alpha)\) percentile interval is
\(\hat{\phi} \pm \tau z_\alpha\) which can also be represented as
\([\hat{\phi}^*_\alpha,\hat{\phi}^*_{1-\alpha}]\). This means that the
interval on the \(\theta\) scale can be defined as
\[\hat{\theta}^*_\alpha = g^{-1}(\hat{\phi} \pm \tau z_\alpha).\] This
also can be represented as an interval,
\[\left [ g^{-1}(\hat{\phi} \pm \tau z_{1-\alpha}), g^{-1}(\hat{\phi} \pm \tau z_\alpha) \right ].\]
Therefore, the percentile method produces a correct interval for
\(\phi\) and, due to the transformation invariance, also produces a
correct percentile interval for \(\theta\). This method assumes the
existence of some monotone normalizing mapping
\(\hat{\phi} = g(\hat{\theta}), \phi = g(\theta)\) and relies on that to
create a correct interval. Since the process is automatic, we do not
need to know the transformation itself, only that it exists. However, in
some cases, no monotone normalizing mapping will exist \citep{Efron86}.

@@

Therefore, we can say that we are 95\% confident that the true
population mean of inches of snow that fall in December is between 13.91
inches and 17.72 inches. Our original histogram with the percentile
confidence interval is shown (Figure 4).

@@

The following method we will be looking at is the bias-corrected
percentile method (BC method) which is an improvement upon the previous
percentile method because we now consider the possibility of bias. It
can be shown that \(\hat{\theta}\) is biased upwards relative to
\(\theta\), which means that the confidence intervals should be adjusted
downwards \citep[\citet{EfronCasi}]{Efron86}. From our simulated
bootstrap replications
\(\hat{\theta}^{*1}, \hat{\theta}^{*2}, \dots ,\hat{\theta}^{*B},\)
define \[p_0 = \frac{\#\{\hat{\theta}^{*b} \leq \theta\}}{B},\] and
define the bias-correction value \[z_0 = \Phi^{-1}(p_0),\] where
\(\Phi^{-1}\) is the inverse function of the standard normal cdf. Thus,
we define a transformation
\(\hat{\phi} = g(\hat{\theta)}, \phi = g(\theta)\) such that for any
\(\theta\), \[\hat{\phi} \sim N(\phi - z_0\tau, \tau^2),\] with \(z_0\)
and \(\tau\) constants. This means that we can say the bias-corrected
method has an \(\alpha\)-level endpoint which can be represented as
\[\hat{\theta}_{BC}[\alpha] = \hat{G}^{-1} \left [ \Phi \left ( 2z_0 + z_\alpha\right ) \right ].\]
If \(\hat{G} = 0.50\), then half of the bootstrap distribution is less
than \(\hat{\theta}\) and our bias-correction value \(z_0 = 0\). In this
case, the confidence interval produced by BC would be the same interval
that was produced by the percentile method.

@@

Using this method to create a 95\% confidence interval has a couple more
steps because we need to find the bias-correction value
\citep{EfronCasi}. First, we calculate the sample mean:

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{alpha\_lower }\OtherTok{\textless{}{-}} \FloatTok{0.025}\NormalTok{; alpha\_upper }\OtherTok{\textless{}{-}} \FloatTok{0.975}

\NormalTok{new\_lower }\OtherTok{\textless{}{-}} \FunctionTok{pnorm}\NormalTok{(}\DecValTok{2}\SpecialCharTok{*}\NormalTok{z0 }\SpecialCharTok{+} \FunctionTok{qnorm}\NormalTok{(alpha\_lower)); }
\NormalTok{new\_upper }\OtherTok{\textless{}{-}} \FunctionTok{pnorm}\NormalTok{(}\DecValTok{2}\SpecialCharTok{*}\NormalTok{z0 }\SpecialCharTok{+} \FunctionTok{qnorm}\NormalTok{(alpha\_upper)); }

\FunctionTok{qdata}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{ mean, }\FunctionTok{c}\NormalTok{(new\_lower, new\_upper), }\AttributeTok{data =}\NormalTok{ dec\_means)}
\end{Highlighting}
\end{Shaded}

@@

Therefore, we are 95\% confident that the true mean inches of snow that
fall in December in Grand Rapids, Michigan is between 14.01 inches and
17.82 inches. Our histogram (Figure 5) is shown.

@@

The bias-corrected and accelerated method (BCa) further modifies the BC
interval. This method does not assume that the standard error, \(\tau\),
is constant as we did in the BC interval
\citep[\citet{EfronCasi}]{Efron86}. Rather, we assume the existence of a
monotone transformation
\(\hat{\phi} = g(\hat{\theta)}, \phi = g(\theta)\) such that for any
\(\theta\),
\[\hat{\phi} \sim N(\phi - z_0\tau_\phi, \tau_\phi^2) \text{ where } \tau_\phi = 1+ a\phi.\]
The \(a\) is known as the acceleration and is a constant that describes
how the standard deviation of \(\hat{\phi}\) varies with \(\phi\). In
other words, \(a\) is proportional to the skewness of the bootstrap
distribution. For example, \(a=z_0\) for one-parameter exponential
families. However, there are many different algorithms to compute and
estimate \(a\) \citep{Flowers18}. Now, our \(\alpha\)-level endpoint for
a BCa confidence interval is
\[\hat{\theta}_{BCa}[\alpha] = \hat{G}^{-1} \left [ \Phi \left ( z_0 + \frac{z_0+z_\alpha}{1-a(z_0+z_a)} \right ) \right ].\]
We can observe that if \(a = 0\), then
\(\hat{\theta}_{BCa}[\alpha] = \hat{\theta}_{BC}[\alpha].\) When
calculating a BCa confidence interval, the acceleration value, \(a\), is
not a function of the bootstrap distribution and must be calculated
separately. However, the process is algorithmic and can be calculated
without too much work. As we saw, each of the three previous methods
(percentile, BC, and BCa) all build upon each other and have less
restrictive assumptions, but computation increases as we loosen
assumptions.

@@

In practice, we can use the jackknife procedure to estimate the
acceleration value for this method \citep{EfronCasi}. The jackknife
procedure is a different resampling technique often used to estimate the
bias and variance of a large population. Thus, we will run the jackknife
procedure:

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{jack\_mean }\OtherTok{\textless{}{-}} \FunctionTok{mean}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{ jackknife\_results}\SpecialCharTok{$}\NormalTok{jack.values); jack\_mean}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{estimated\_a }\OtherTok{\textless{}{-}}\NormalTok{ (}\DecValTok{1}\SpecialCharTok{/}\DecValTok{6}\NormalTok{)}\SpecialCharTok{*}\FunctionTok{sum}\NormalTok{((jackknife\_results}\SpecialCharTok{$}\NormalTok{jack.values }\SpecialCharTok{{-}}\NormalTok{ jack\_mean)}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{)}\SpecialCharTok{/}
\NormalTok{  (}\FunctionTok{sum}\NormalTok{((jackknife\_results}\SpecialCharTok{$}\NormalTok{jack.values }\SpecialCharTok{{-}}\NormalTok{ jack\_mean)}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{))}\SpecialCharTok{\^{}}\NormalTok{(}\FloatTok{1.5}\NormalTok{); estimated\_a}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{a\_new\_lower }\OtherTok{\textless{}{-}} \FunctionTok{pnorm}\NormalTok{(z0 }\SpecialCharTok{+}\NormalTok{ (z0 }\SpecialCharTok{+} \FunctionTok{qnorm}\NormalTok{(alpha\_lower))}\SpecialCharTok{/}
\NormalTok{                    (}\DecValTok{1}\SpecialCharTok{{-}}\NormalTok{estimated\_a}\SpecialCharTok{*}\NormalTok{(z0 }\SpecialCharTok{+} \FunctionTok{qnorm}\NormalTok{(alpha\_lower)))); }
\NormalTok{a\_new\_upper }\OtherTok{\textless{}{-}} \FunctionTok{pnorm}\NormalTok{(z0 }\SpecialCharTok{+}\NormalTok{ (z0 }\SpecialCharTok{+} \FunctionTok{qnorm}\NormalTok{(alpha\_upper))}\SpecialCharTok{/}
\NormalTok{                     (}\DecValTok{1}\SpecialCharTok{{-}}\NormalTok{estimated\_a}\SpecialCharTok{*}\NormalTok{(z0 }\SpecialCharTok{+} \FunctionTok{qnorm}\NormalTok{(alpha\_upper)))); }
\FunctionTok{qdata}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{ mean, }\FunctionTok{c}\NormalTok{(a\_new\_lower, a\_new\_upper), }\AttributeTok{data =}\NormalTok{ dec\_means)}
\end{Highlighting}
\end{Shaded}

@@

The final method we explain is the bootstrap-t interval (or the
studentized method). Recall from earlier, our sample, \(X\) from which
we can calculate an estimate (\(\hat{\theta}(X)\)) of the parameter of
interest \(\theta\) \citep[\citet{Puth15}]{Efron86}. We can also
estimate \(\hat{\sigma}(X)\) for the standard error of \(\theta\). We
can use these parameters to find the Student's \(t\)-statistic defined
as: \[T = \frac{\hat{\theta} - \theta}{\hat{\sigma}}.\] As such, the
\(\alpha\)th percentile point of a confidence interval of \(\theta\)
would be \(\hat{\theta} - \hat{\sigma}T_{(\alpha)}\) where
\(T_{(\alpha)}\) represents the \(\alpha\)th percentile of the
\(t\)-distribution, \(T\). Unfortunately, the percentiles of this
\(t\)-distribution are unknown in most cases, but we can use
bootstrapping to estimate these percentiles. To do this, we perform a
large number, \(B\), of bootstrap samples, from which we can find the
bootstrap replications of the parameter of interest,
\(\hat{\theta^*} = \hat{\theta}(X^*)\), and the standard error,
\(\hat{\sigma^*} = \hat{\sigma}(X^*)\). From these, we can calculate a
t-statistic for each bootstrap sample:
\[T^* = \frac{\hat{\theta}^* - \hat{\theta}}{\hat{\sigma}^*}.\] Using a
large number of these bootstrap samples, we can estimate the percentiles
of the \(t\)-distribution such that:
\[\hat{T}_{(\alpha)} = B*\alpha\text{th ordered value of all the bootstrap replications of }T^*.\]
This means that for \(B = 2,000\) and \(\alpha = 0.90\), then
\(\hat{T}_{(\alpha)}\) is the 1,800th ordered point of all of the
bootstrap replications of \(T^*\). It follows that the \(\alpha\)th
studentized confidence interval endpoint can be given with
\[\hat{\theta}_T[\alpha] = \hat{\theta} - \hat{\sigma}T_{(\alpha)},\]
where we can estimate the standard error using
\[\hat{\sigma} = \frac{1-\hat{\theta}^2}{\sqrt{n}}.\] One of the main
factors why the studentized method is so popular is that we assume that
our bootstrapped statistic is pivotal, which means that the confidence
interval does not depend on any other parameters. Instead, we can
calculate the appropriate confidence interval for the parameter of
interest specifically from the bootstrapped statistics
\citep[\citet{Puth15}]{Efron86}.

@@

So we are 95\% confident that the true mean number of inches of snow
that fall in December in Grand Rapids, Michigan is between 14.03 inches
and 18.02 inches. This would give us the following confidence interval
(Figure 7):

@@

\includegraphics{paper_files/figure-latex/unnamed-chunk-22-1.pdf}

As shown in Figure 10, the studentized method had the greatest accuracy
in creating the confidence intervals that captured the true mean. At
small sample sizes such as 5, the studentized method has an accuracy of
82.3\%, while the following best is the BCa method with a 73.4\%
accuracy. As the sample size increases, all methods' accuracy increases
quickly and performs comparably, although the studentized and BCa
methods perform slightly better. All the intervals stay constant at
around 92.5\% to 93.5\% accuracy for large sample sizes. Again, the
proportions are shown (Table 3) for each sample size for reference.
